Asymmetric relationships and their effects on coexistence

### 2023-03-23 paper/Paper.Rnw
pretty complete job editing for now, just waiting for Jasmin's figure work to complete my editing

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships - between organisms and between organisms and their environments. 
Ecologists may ask, for instance, how fluctuations in local- and regional-scale environments relate to 
population dynamics. To study such relationships and to understand how two variables interact, ecologists 
very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
and direction of association between variables often don't tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what we have previously referred to as 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
regression), but correlations are extremely common, and other methods also reveal only a small portion 
of the information content of the relationship between variables \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recently several 
insights have been gained by studying how features of associations beyond standard measures 
impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_AER} show that ATAs are common in ecology
and can influence Taylor's law and Moran effects; \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
ecosystem functioning. Fig. \ref{fig:pedagog1}g,h show contrasting examples of ATAs in nature using 
plankton population density time series. To our knowledge, the implications of ATAs for competition and coexistence have
not yet been investigated. Doing so is the overarching goal of this study.

@@

%Intro to storage effects and why they might be influenced by ATAs
It is possible to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
may be relevant to them. Storage effects can allow for species to coexist while fluctuating in abundance through time
because storage effects can provide a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
Storage effects hinge on positive covariation between the fluctuating quality of the environment
for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
and on the assumption that optimal environmental conditions for distinct species are also distinct,
so that no two species find the same conditions to be optimal. Then, good environmental
conditions for a currently common species are paired with strong intraspecific competition, because of
positive EC covariance and the commonness of the species. But good environmental conditions for 
a currently rare species are not as closely accompanied by strong competition because the rarity of
the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model is as follows.
Letting $N_i(t)$ denote the population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is a mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that individuals die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).

%***DAN: I commented the below because we decided it does not belong in the Intro, but a version of it 
%should be in the Discussion. Keeping it for now until I have a chance to move it there. Note also that
%I now think this material needs some editing for clarity, some of that was done in a paper version I 
%marked up in January 2023.
% %Climate change may alter ATAs of coexistence-relevant ATAs
% An additional motivation for quantifying the importance of ATAs for species coexistence is that there is reason to believe climate change may
% alter ATAs of environmental variables which impact coexistence, as follows.
% It is well known that climate change can alter the intensity, duration, extent and frequency of extreme environmental events, including
% heat waves, floods, hurricanes and fires \citep{Meehl_2004, Jentsch_2007, Ummenhofer_2017, Lyon2019, Keelings2020}.  
% If extreme events are also longer lasting, then the nature of the covariation between phenologically separated environmental signals 
% may be modified by changes in extreme climatic events; in particular, ATAs between such variables may be altered. 
% For instance, imagine a case in which early spring precipitation influences the growth of a plant species, A,
% whereas late spring precipitation influences plant species B. If these species are competitors, then EC covariation
% contributing to their coexistence, or failure to coexist, can boil down to covariation 
% between early and late spring precipitation. Because longer lasting weather events are more likely to contribute
% to both early and late spring rainfall variables, a tendency for spring rainfall events to become both more extreme and 
% longer lasting increases the degree of right-tail association between early and late spring rainfall. 
% Likewise, a tendency for spring droughts to become both more intense and longer lasting can increase
% left-tail association between these variables. As we will show in this study, changes in ATAs can then
% result in different competitive outcomes between species. 
% \cite{Chesson_2012} provides examples from an Australian grass-forb community and from an Arizona 
% community of desert annuals that seem to have all the ingredients necessary to make effects of the
% type we describe here a possibility.
% To our knowledge, the potential importance of climatic changes in ATAs for species coexistence has 
% not been studied, but the importance for coexistence of changes 
% in means and variability of environmental variables has been studied in many systems
% [e.g., \cite{White_2001, Adler_2006, Jentsch_2007, Angert_2009, diPaola_2012, Descombes_2020, Usinowicz_2021}].
% Changes in variability and in extreme events are distinct from the changes in ATAs we 
% consider here, because ATAs pertain to \emph{relationships} between environmental and other
% variables in the extremes, which are distinct from and can be altered independently
% of changes in the extremes of the univariate marginal distributions themselves \citep{Nelsen2006,Ghosh_2020_AER}.

@@

We develop our new theory for a general two-species model, later applying it to specific scenarios.
Our mathematical notation parallels
that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$, i.e., $r_i(E_i,C_i)$. We also assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

@@

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is then determined in one of 
four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.

@@

The other mechanisms in MCT that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas ATAs are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
coexistence operating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of MCT that was goal G1 of the Introduction.

@@

%Setup - two versions, and basics which apply to both
To pursue goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
comparing the magnitude of the ATA effect on coexistance, $\Delta_i^{[EC]}$, to the magnitudes 
of the other terms of 
the decomposition (\ref{eq:full_decomp}) for various model parameterizations. We used two versions 
of the lottery model which we now specify. The versions both used the same dynamical equations
(Introduction) and assumed temporally independent and identically distributed (iid) fecundities, but the
versions differed in the distributions of $(B_1,B_2)$ used. Both distributions were based on 
transformations of the distributions $(b_1,b_2)$ pictured in Fig. \ref{fig:pedagog1}a-c;
the details of how the left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c), 
and symmetrically (Fig. \ref{fig:pedagog1}b) tail associated alternatives for $(b_1,b_2)$ were generated 
are in SI section \ref{SI-sect:noise}.

@@

%G2 
To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
of the contributions to coexistence that come from ATAs compared to other mechanisms; and to see if ATA rescue 
or ATA exclusion can occur), Fig.
\ref{fig:fig3} shows that, for the log-normal fecundities lottery model, 
the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
$\mu_2$, i.e., for closer-to-equal competitors. 
In the $\mu_1=\mu_2$ case (panels a-c), for which the two species were equal 
competitors, storage effects ($\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$) 
were the only means by which coexistence could occur, and ATA effects
tended to be negative, weakening total storage effects.

@@

%G3 ***DAN: check this again after Jasmin gets done making the new version of Fig. 5
Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition, and that 
ATA rescue occurs for some combinations of parameters for this system (Fig. \ref{fig:fig5}a-c).

@@

%Summary-type paragraph of main results
We produced and applied a new extension of MCT that 
quantifies the impact of ATAs on species coexistence and the maintenance of biodiversity. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to models based on a laboratory 
system of competing diatoms demonstrated that  
ATA contributions to storage effects and to species coexistence and the maintenance of biodiversity
can often be comparable in magnitude to the contributions of previously recognized mechanisms.
Our results highlight that common methods of correlation and covariance fail to include 
important information on the structure of relationships between variables. Our results add
coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined to constitute the 
contribution of \emph{associations} between $E$ and $C$, including both ATA effects and
correlation \emph{per se}; the prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association,
such as ATAs, which we now know are important.

%ATA specialists
Contributions of ATAs to species coexistence open the possibility that
``ATA specialists'' may exist in some contexts. If a strong competitor is present across a
collection of habitat patches and the GWR of a weaker competitor is only positive 
in a patch in the presence of 
ATAs between the temporal fluctuations of two environmental variables in that patch,
then the weaker competitor can only persist in patches for which ATAs are present.
The weaker competitor can therefore perhaps be regarded as an ``ATA specialist'' in the landscape. EC-associations
can often boil down to associations between two environmental variables, as was assumed to be the case
for the lottery model. Mechanisms of coexistence should correspond to
means by which niche differentiation is possible. Thinking about our new mechanism of 
coexistence in terms of possible ATA specialists helps highlight this connection. Likewise,
because we have described ATA exclusion as well as ATA rescue, it is possible to 
imagine a scenario in which a weaker competitor can only exist in habitat patches 
for which ATAs are absent, and is hence a specialist on scenarios with symmetric
tail associations of environmental variables (an ``STA specialist''). Additional 
research will be needed to assess to what degree these possibilities are realized 
in nature.

%future work: study ATAs in environmental variables and how theyre changing - tie in with specialist
In the Introduction we explained a plausible means by which climate change may
alter the tail association properties of phenologically separated environmental variables
such as early- and late-spring rainfall. The possibility 
that changes in the nature of environmental ATAs 
may result in the extirpation of ATA and STA specialists indicates that it may
be important to engage in further study of ATAs between environmental variables 
and how they differ across space and may be changing through time. 
Extreme environmental events have been much studied, including
how the strength and commonness of such events may be changing with climate change, and how such changes 
may influence coexistence (see citations in the Introduction). But
studies of changes in environmental ATAs would differ from such prior work because
ATAs pertain to \emph{relationships} between variables in the extremes.

>>>>

%Intro the paper by intro'ing ATAs
At the core of ecology is the study of relationships - between organisms and between organisms and their environments. 
Ecologists may ask, for instance, how local environment relates to 
population growth rate. To study such relationships and to understand how two variables interact, ecologists 
very commonly use metrics such as correlation and covariance. But these standard measures of the strength 
and direction of association between variables often do not tell the whole story 
\citep{Ghosh_2020_AER}.
To illustrate, the three pairs of variables in Fig. \ref{fig:pedagog1}a-c all have the same 
Pearson correlation despite striking differences in association structure: Fig. \ref{fig:pedagog1}b shows a 
symmetric type of association, while Fig. \ref{fig:pedagog1}a,c exhibit what we have previously referred to as 
``asymmetric tail association'' [ATA; \citealp{Ghosh_2020_AER}]. 
Metrics of association other than correlations and covariance are also often used (e.g., various forms of 
regression), but correlations are extremely common, and other common methods also reveal only a small portion 
of the information content of the relationship between variables \citep{Nelsen2006}.
ATAs and related ideas had previously been seldom considered in ecology, but recently several 
insights have been gained by studying how features of associations beyond standard measures 
impact various ecological phenomena \citep{deValpine_2014, Anderson_2019, Popovic_2019, Ghosh_2020_AER, Ghosh_2020_Eco,
Ghosh_2020_ecolevol, Ghosh_2021}. For instance, \cite{Ghosh_2020_AER} show that ATAs are common in ecology
and can influence Taylor's law and Moran effects; \cite{Ghosh_2020_Eco} argue that extinction risks can become elevated 
when metapopulations are subject to ATAs; and \cite{Ghosh_2021} show that ATAs can alter the stability of 
ecosystem functioning. Fig. \ref{fig:pedagog1}g,h show contrasting examples of ATAs in nature using 
plankton population density time series. To our knowledge, the implications of ATAs for competition and coexistence have
not yet been investigated. Doing so is the overarching goal of this study.

@@

%Intro to storage effects and why they might be influenced by ATAs
It is possible to see, intuitively, 
why ATAs may influence one mechanism of coexistence in particular, storage effects. We review, conceptually, 
what storage effects are \citep{Chesson_1994,Chesson_2012,Ellner_2016,Barabas_2018}, and explain why ATAs 
may be relevant to them. Storage effects allow competing species to fluctuate in abundance without any going extinct
by providing a relative benefit to whichever species is currently rare \citep{Chesson_2000}.
Storage effects hinge on positive covariation between the fluctuating quality of the environment
for a species (E) and the strength of competition (C) experienced by that species (called EC covariance);
and on the assumption that optimal environmental conditions for distinct species are also distinct,
so that no two species find the same conditions to be optimal. Then, good environmental
conditions for a currently common species are paired with strong intraspecific competition, because of
positive EC covariance and the commonness of the species. But good environmental conditions for 
a currently rare species are not as closely accompanied by strong competition because the rarity of
the species limits competitive influence. The resulting asymmetry is what provides the relative benefit to the
rare species. Storage effects can be quantified by measuring to what extent EC covariance contributes to 
differences between currently rare and currently common species growth rates \citep{Chesson_1994,Ellner_2016}.  
ATAs may, in principle, influence storage effects simply because storage effects hinge on EC covariation, 
and ATAs, when they occur between environment and competition variables, can be an aspect of EC covariation.

%Lottery model and results based on it
To further demonstrate the plausibility of ATAs altering coexistence between species, we carried out 
simple simulations using a two-species model, the well-known lottery model. The simulations described in this
paragraph are sufficient to demonstrate that ATAs influence coexistence in some manner; the
precise nature of that influence is the subject of the rest of the paper.
The lottery model, which was originally introduced as a model of coral reef fish competing for a set of $N$ breeding territories, 
is as follows.
Letting $N_i(t)$ denote the adult population density of species $i=1,2$ at time $t$, and defining $N = N_1(t)+N_2(t)$,
model equations are
\begin{equation}
N_i(t+1)=(1-\delta)N_i(t)+\delta N \frac{B_i(t)N_i(t)}{B_1(t)N_1(t)+B_2(t)N_2(t)}  \label{model_eq}
\end{equation}
for $i=1,2$. Here, $\delta$ is an adult mortality rate, and $B_i(t)$ is the fecundity of species $i$ at time $t$. 
The model postulates that adults die at rate $\delta$ at each time step, and are replaced by juveniles 
in proportion to the reproductive outputs of the two species that year. Note that $N$ is constant through time. 
Notation for the whole paper is summarized in Table \ref{tab:notation}.
For each $i$, we assume for simplicity that the random variables $B_i(t)$ are independent and identically 
distributed (iid) through time. In three distinct similations (Fig. \ref{fig:pedagog1}d-f), 
$B_1$ and $B_2$ were, respectively, 
left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c) and symetrically (Fig. \ref{fig:pedagog1}b)
tail associated, while the same marginal distributions for these fecundities were used in all simulations (see Methods
for details). Thus, the model setup was identical in the three simulations except for different ATAs. 
ATAs substantially influenced aspects of species coexistence in these simulations (Fig. \ref{fig:pedagog1}d-f).

@@

We develop our new theory for a general two-species model, later applying it to specific scenarios.
Our mathematical notation parallels
that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$, i.e., $r_i(E_i,C_i)$. We also assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though the iid assumption is not necessary, it simplifies the
presentation of new ideas.

@@

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is then determined in one of 
four alternative ways: either by setting $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.

@@

The other mechanisms in MCT that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas ATAs are 
strictly about covariation. Therefore, storage effects are the only mechanism of coexistence
for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
coexistence operating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of MCT that was goal G1 of the Introduction.

@@

%Setup - two versions, and basics which apply to both
To achieve goal G2 of the Introduction, we applied the framework of Theory to the lottery model, 
comparing the magnitude of the ATA effect on coexistance, $\Delta_i^{[EC]}$, to the magnitudes 
of the other terms of 
the decomposition (\ref{eq:full_decomp}) for various model parameterizations. We used two versions 
of the lottery model which we now specify. The versions both used the same dynamical equations
(Introduction) and assumed temporally independent and identically distributed (iid) fecundities, but the
versions differed in the distributions of $(B_1,B_2)$ used. Both distributions were based on 
transformations of the distributions $(b_1,b_2)$ pictured in Fig. \ref{fig:pedagog1}a-c;
the details of how the left- (Fig. \ref{fig:pedagog1}a), right- (Fig. \ref{fig:pedagog1}c), 
and symmetrically (Fig. \ref{fig:pedagog1}b) tail associated alternatives for $(b_1,b_2)$ were generated 
are in SI section \ref{SI-sect:noise}.

@@

\subsection{Goal G2, lottery model results}

To begin fulfilling goal G2 of the Introduction (to explore, using the lottery model, the relative magnitudes 
of the contributions to coexistence that come from ATAs compared to other mechanisms; and to see if ATA rescue 
or ATA exclusion can occur), Fig.
\ref{fig:fig3} shows that, for the log-normal fecundities lottery model, 
the ATA term of (\ref{eq:full_decomp}) sometimes, but not always, has 
magnitude comparable to the magnitudes of other terms of the extended MCT decomposition. 
Magnitudes of $\Delta_i^{[EC]}$ tended to be larger, relative to the magnitudes of the 
other terms of (\ref{eq:full_decomp}), for smaller-magnitude differences between $\mu_1$ and 
$\mu_2$, i.e., for closer-to-equal competitors. 
In the $\mu_1=\mu_2$ case (panels a-c), for which the two species were equal 
competitors, storage effects ($\Delta_i^{[EC]}+\Delta_i^{[E\nat C]}$) 
were the only means by which coexistence could occur, and ATA effects
tended to be negative, weakening total storage effects.

@@

\subsection{Goal G3, results for the diatom system}

%***DAN: check this again after Jasmin gets done making the new version of Fig. 5
Starting to fulfill goal G3 of the Introduction (to provide an empirical example of species coexistence 
which highlights the role ATAs can play), Fig. \ref{fig:fig5}a-c 
show that ATA contributions to coexistence often have magnitudes comparable to the 
magnitudes of other terms of the extended MCT decomposition, and that 
ATA rescue occurs for some combinations of parameters for this system (Fig. \ref{fig:fig5}a-c).

@@

%Summary-type paragraph of main results
We produced and applied a new extension of modern coexistence theory (MCT) that 
quantifies the impact of asymmetric tail associations (ATAs) on species coexistence and the maintenance of biodiversity. 
Building upon a recent, simulation-based approach to MCT \citep{Ellner_2016, Ellner_2019},
we decomposed storage effects into two new mechanisms, the contribution of ATAs to coexistence,
and the contribution of $EC$ correlation \emph{per se}.
Applications of our extended theory to the lottery model and to a laboratory diatom system demonstrated that  
ATA contributions to storage effects and to species coexistence 
can often be comparable in magnitude to the contributions of previously recognized mechanisms.
Our results add
coexistence and biodiversity maintenance to the list of ecological theories and phenomena on which 
ATAs have been shown to have an important impact \citep{Ghosh_2020_AER, Ghosh_2020_Eco, Ghosh_2020_ecolevol, Ghosh_2021}.  
Though storage effects are typically described as the contribution
of $EC$ \emph{covariance} to coexistence \citep{Chesson_2012, Ellner_2016, Ellner_2019},
our results suggest that storage effects should be redefined to constitute the 
contribution of \emph{associations} between $E$ and $C$, generally, including both ATA effects and
correlation \emph{per se}. The prior terminology, \emph{covariance}, is typically construed 
as reflecting the linear relationship between variables, ignoring complexities of the association,
such as ATAs, which we now know can be important. ATAs contributions to coexistence were less important than other mechanisms 
in some contexts (the log-normal fecundities lottery model), but were very important in other contexts
(the beta fecundities lottery model), including our empirically
based example (the diatom system). Though future work should seek to understand precisely when ATAs are or are not important
for coexistence, our work demonstrates the overall importance of this new mechanism.

%ATA specialists
Contributions of ATAs to species coexistence open the possibility that
``ATA specialists'' may exist. For instance, if a strong competitor is present across a
collection of habitat patches on a landscape and the GWR of a weaker competitor is only positive 
in a patch if there are 
ATAs between the temporal fluctuations of two environmental variables in that patch,
then the weaker competitor can only persist in patches for which ATAs are present.
The weaker competitor can therefore perhaps be regarded as an ``ATA specialist'' in the landscape. EC-associations
can boil down to associations between two environmental variables, as was the case
for the lottery model. For the lottery model, $B_i(t)=E_i(t)$ was assumed, and $C_{i \bs i}$ simplified to $E_j(t)/\delta$ (see SI section 
\ref{SI-SIsect:TheoryForLotteryBoth}). Thus, EC-associations parallel associations between the two
environmental variables $E_i(t)$ and $E_j(t)$ for the lottery model, and ATA specialists may exist. 
Mechanisms of coexistence should correspond to
means by which niche differentiation is possible. Thinking about our new mechanism of 
coexistence in terms of possible ATA specialists helps highlight this connection. Likewise,
because we have described ATA exclusion as well as ATA rescue, it is possible to 
imagine a scenario in which a weaker competitor can only exist in habitat patches 
for which ATAs between environmental variables are absent, and is hence a specialist on scenarios with symmetric
tail associations of variables (an ``STA specialist''). Additional 
research will be needed to assess to what degree these possibilities are realized 
in nature.
